---
title: "Statistical and econometric data analysis with R "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Do more frequent medical consultations promote better well-being?

- This project aims to investigate whether more annual visits among older people relate to better well-being and health.

- Project was prepared as a project for university classes of  Data Visualization techniques. 

- The study is based on data from the seventh wave of the Panel Study of Health, Population Aging and Retirement Processes "SHARE: 50+ in Europe and covers Poland, Sweden and Slovakia. 

- As a result of a series of statistical tests and econometric modeling, it turns out that the number of medical consultations is a statistically significant factor in the well-being of people aged 50+, but only for Poland and Sweden. 

- Moreover, the relationship between this indicator and health status is inverted: a higher number of visits does not at all whole promote better well-being. This confirms the existence of non-objective health reasons for the high number of annual visits among the elderly.

#### Methodology 

- In 2018, according to an OECD report, the average number of medical consultations per person was 6.7 visits in the European Union. In Poland, this figure was slightly higher at 7.6 visits. Slovakia had the highest average number of visits, and Sweden the lowest. Hence the choice of countries for the analysis: so it was decided to study how better well-being is affected by frequent, average and infrequent visits to the doctor 

- The determinant of well-being was chosen to be the maxgrid variable, which represents handshake strength. This measure is used to assess muscle
strength primarily among the elderly. Numerous studies indicate that handshake strength is a biomarker of health not only physical, but also
mental, and overall well-being, as it is correlated with the state of the muscular, circulatory, digestive, skeletal, nervous systems of the
broader quality of life

- The SHARE survey includes people aged 50+ and, in addition, also collects information on the partners of the subjects, so the study sample was
limited to people over 50 only. In addition, in order to avoid presenting outlier-distorted results, observations in which the explanatory
variable, handshake strength index, and frequency of visits to the doctor exceed 1.5 quartile divergence lengths were removed from the sample. In
addition, observations for which these variables were coded as negative (implying missing data) were removed. The same was done for all control
variables that entered the regression model

- The significance of differences between Poland, Sweden and Slovakia in terms of the frequency of medical consultations and levels of well-being
was tested using a weighted one-way ANOVA and the non-parametric Kruskal-Wallis test. 

- Additionally, weighted Student's t-tests were conducted to determine which countries in particular differ significantly. 

- To test the fulfillment of the assumptions of ANOVA analysis, the Levene test for homogeneity of variance was conducted, and the normality of
the distribution of variables was checked using the Kolmogorov-Smirnov test.  

- The relationship between well-being and frequent visits to each study country was additionally tested using the chi-square test, and its
strength was measured by the V-Cramer coefficient. 

- To enable the chi-square test of concordance, the continuous study variables were aggregated into categorical variables.

- The determinants of handshake strength, along with the average number of visits to the doctor, were determined by constructing three
multivariate linear regressions for Poland, Sweden and Slovakia. 

- All descriptive statistics and statistical tests were calculated on the entire population by using appropriate weights.

- All decisions regarding the absence of reasons to reject or reject the null hypotheses are made at the 0.05 significance level. 

## Set up
Activate all packages (previously they could be installed with install.packages('')).
```{r message=FALSE, warning=FALSE}
library(foreign)
library(ggplot2)
library(sjstats)
library(Hmisc)
library(dplyr)
library(plyr)
library(GGally, quietly = TRUE)
library(DescTools)
library(car)
library(poliscidata)
library(stargazer)
```


## Data preparation 

```{r}
# Clean environment 
rm(list=ls())
```


```{r, echo=FALSE}
setwd("/Users/ovtsho/Documents/SGH/SGH V sem/Wizualizacja/dane")
load("easySHARE_rel7_0_0.rda")
```


```{r}

# Set working directory and load data 

SHARE.w7 <- dplyr::filter(easySHARE_rel7_0_0, wave==7)
SHARE.weights <- read.dta("sharew7_rel7-0-0_gv_weights.dta")
SHARE.weights <- dplyr::select(SHARE.weights, mergeid, cciw_w7)
SHARE <- dplyr::left_join(SHARE.w7, SHARE.weights, by="mergeid")
attach(SHARE) 
rm(easySHARE_rel7_0_0, SHARE.w7, SHARE.weights)
SHARE$country.lab <- substr(mergeid,1,2)

SHARE$hc002_mod[SHARE$hc002_mod < 0] <- NA
SHARE$maxgrip[SHARE$maxgrip < 0] <- NA

SHARE.PL <- dplyr::filter(SHARE, age >= 50 & country.lab=="PL" & !is.na(cciw_w7))
SHARE.SE <- dplyr::filter(SHARE, age >= 50 & country.lab=="SE" & !is.na(cciw_w7))
SHARE.SK <- dplyr::filter(SHARE, age >= 50 &country.lab=="SK" & !is.na(cciw_w7))

SHARE <- dplyr::filter(SHARE, age >= 50 & country.lab %in% c("PL", "SE", "SK") & !is.na(cciw_w7))

SHARE$country.lab <- as.factor(SHARE$country.lab)
levels(SHARE$country.lab)[levels(SHARE$country.lab)=='PL'] <- 'Poland'
levels(SHARE$country.lab)[levels(SHARE$country.lab)=='SE'] <- 'Sweden'
levels(SHARE$country.lab)[levels(SHARE$country.lab)=='SK'] <- 'Slovakia'
```


## Basic descriptive statisitcs for samples 

- The largest group of respondents among the three selected countries is Poland, with more than 4,000 respondents, and the smallest is Slovakia, with 2,000. 

- The gender structure in the sample of the three countries maintains the standard split between a larger female group (52%-55%) and a smaller
male group (45-47.5%).

- Respondents from Sweden (72.15) have the highest average age (shown in figure  with a red dot) among the sample, and the youngest on average
are respondents from Slovakia (62.54). Respondents from Poland were on average 66 years old. 

- From the graph of the sample's age structure, it can be inferred that Sweden's sample is older on average, as it includes more people aged 70+
and has the highest median age among the three groups. 

- It is interesting to note that Slovakia, with the highest average frequency of visits to the doctor (according to OECD data), is the youngest sample, dominated by people under 70.  

```{r}
SHARE.m<-ddply(SHARE, .(country.lab), summarize, 
              mean=round(mean(age, na.rm=TRUE), 2))
ggplot(SHARE, aes(x=country.lab, y=age, fill=country.lab))+
  geom_violin()+
  geom_boxplot(width=0.3, alpha=0.2, outlier.shape = NA)+
  geom_point(data=SHARE.m, aes(x=country.lab, y=mean, color='red', size=4))+
  theme_bw()+
  labs(title = "Sample age structure")+
  ylab("Age")+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        text = element_text(size = 16))+
  scale_fill_brewer(palette="PuBuGn")


SHARE$female <- as.factor(SHARE$female)
levels(SHARE$female)[levels(SHARE$female)==1] <- 'Female'
levels(SHARE$female)[levels(SHARE$female)==0] <- 'Male'

group.colors <- c(Female = "#FB8281", Male = "#A4DDEE")

options(scipen=999) 
ggplot(SHARE, aes(x=country.lab, fill=female))+
  geom_bar(width = 0.7)+
  theme_bw()+ylab('Sample size')+
  labs(fill='Sex',title = "Sample size by gender structure")+
  geom_text(aes(by=country.lab), stat="prop",  position = position_stack(vjust = .5), size=5)+
  theme(axis.title.x=element_blank(),
        legend.position = c(.8, .85),
        text = element_text(size = 16))+ 
  scale_fill_manual(values=group.colors)
```



## Analysis outcomes 

- Data from the seventh wave of the SHARE survey shows that respondents from Poland had the highest average number of medical consultations, with
almost 6 visits per year, but this value is quite variable, as the number of annual visits deviated from the average by almost 5 visits on
average. 

- All three selected countries  have fairly wide confidence intervals for the weighted average number of visits to a doctor. 

- Sweden had the lowest number of visits, with an average of 3.5, and in the middle is Slovakia with an average of 4.5

```{r message=FALSE, warning=FALSE}
#Remove outliners with weights
Q1 <- wtd.quantile(SHARE$hc002_mod, .25, na.rm=TRUE, weight=SHARE$cciw_w7)
Q3 <- wtd.quantile(SHARE$hc002_mod, .75, na.rm=TRUE, weight=SHARE$cciw_w7)
IQR <- IQRw(SHARE$hc002_mod, na.rm=TRUE, weight=SHARE$cciw_w7)
SHARE_out <- subset(SHARE, SHARE$hc002_mod > (Q1 - 1.5*IQR) & SHARE$hc002_mod < (Q3 + 1.5*IQR))

```


```{r message=FALSE, warning=FALSE}
data1 <- ddply(SHARE_out, .(country.lab), summarize, 
      wgt.ave = weighted.mean(hc002_mod, cciw_w7, na.rm=TRUE), 
      weighted_var = wtd.var(hc002_mod, cciw_w7, na.rm=TRUE),
      sd=sqrt(weighted_var))

ggplot(data1, aes(country.lab, wgt.ave, fill=country.lab)) + 
  geom_bar(stat="identity", width = 0.7)+scale_fill_brewer(palette="PuBuGn")+
  geom_errorbar(aes(x=country.lab, ymin= wgt.ave-sd, ymax=wgt.ave+sd), width=0.4, colour="orange", alpha=0.9, size=1.3,)+
  theme_bw()+
  labs(fill='', title='Weighted average number of medical consultations', subtitle='per year with confidence interval')+
  ylab('Average number of medical consultations')+
  theme(legend.position="", 
        axis.title.x=element_blank(),
        text = element_text(size = 16))
```

- For a more accurate comparison of the frequency of medical consultations among the countries studied, this variable was aggregated to short
ranges. 

- People most often go for medical consultations up to 3 times a year in Poland, Sweden and Slovakia. 

- It is noteworthy that a larger proportion of respondents from Sweden and Slovakia are clustered in the 0-3 range, while a larger proportion of
respondents from Poland are scattered in the 0-3, 4-6,7-9 and 10-12 ranges, indicating on average a larger number of people who can be classified
as frequent attenders in Poland.   

```{r message=FALSE, warning=FALSE}
# Prepare categorical varibale 
SHARE_out$hc002_mod_cat <- cut(SHARE_out$hc002_mod,
                               breaks=c(0, 3, 6, 9, 12, 15, 20),include.lowest= TRUE, 
                               labels=c('0-3', '4-6', '7-9', '10-12', '13-15', '16-20'))


ggplot(data = SHARE_out, aes(x = hc002_mod_cat, fill=country.lab,  weight=cciw_w7)) +
  geom_histogram(stat="count", position='dodge')+
  scale_y_continuous(breaks=c(0, 2000000, 4000000), labels=c('0', '2 mln', '4 mln'))+
  scale_fill_brewer(palette="PuBuGn")+theme_bw()+
  geom_text(aes(by=country.lab), stat="prop", size = 3, vjust = -1,  position = position_dodge(1.2))+
  labs(fill='Countries', title = 'Weighted average of the number of medical consultations', subtitle='per year')+
  ylab('Population')+
  xlab('Number of medical consultations')+
  theme(panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(),
      legend.position = c(.82, .85), 
      text = element_text(size = 14))
```


- As for handshake strength, which is taken as a measure of well-being among people 50+ in this study, the Swedes have the highest average
strength (34.03) and the Slovaks the lowest (31.43), which is an interesting observation since the Swedish population is more advanced in age
which would have to translate into lower average handshake strength. 

- This phenomenon can be explained by the higher value of the average health expectancy index for Sweden compared to Slovakia and Poland.  

```{r message=FALSE, warning=FALSE}
#Remove outliners for maxgrip variable 
Q1 <- wtd.quantile(SHARE_out$maxgrip, .25, na.rm=TRUE, weight=SHARE_out$cciw_w7)
Q3 <- wtd.quantile(SHARE_out$maxgrip, .75, na.rm=TRUE, weight=SHARE_out$cciw_w7)
IQR <- IQRw(SHARE_out$maxgrip, na.rm=TRUE, weight=SHARE_out$cciw_w7)
SHARE_out <- subset(SHARE_out, SHARE_out$maxgrip > (Q1 - 1.5*IQR) & SHARE_out$maxgrip < (Q3 + 1.5*IQR))


SHARE_out.wm_grip<-ddply(SHARE_out, .(country.lab), summarize, 
                wmean=round(wtd.mean(maxgrip , cciw_w7, na.rm=TRUE), 2))

ggplot(SHARE_out, aes(x=country.lab, y=maxgrip, fill=country.lab))+
  geom_violin(alpha=.9)+
  geom_boxplot(width=0.3, alpha=0.2, outlier.shape = NA, aes(weight=cciw_w7))+ 
  geom_point(data=SHARE_out.wm_grip, aes(x=country.lab, y=wmean, color='red', size=3.5))+
  theme_bw()+
  labs(title='Distribution of population handshake strength')+
  ylab("Handshake power")+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        text = element_text(size = 16))+
  scale_fill_brewer(palette="PuBuGn")
```


## Comparison of results between countries and specific groups 

- A weighted OneWayAnnova test, or one-way analysis of variance - a statistical test used to compare means across multiple populations - was
chosen to examine differences in well-being (handshake strength) between Poland, Sweden and Slovakia. 

- The null hypothesis of equality of means was rejected (p-value=0.00) but the results of this test can be trusted provided the assumptions of
the test are met: similar variance in the studied groups and normal distribution of the studied variables. 

- The homogeneity of the variance was checked using the Levene test, which was not rejected at the 0.05 significance level (p-value= 0.79).

- Normality of the distribution was checked with the Kolmogorov-Smirnov test and was rejected at any significance level (p-value= 0.00). 

- Although, in this study we are dealing with a large sample (more than 9,000 observations), so the limiting distribution of the studied
variables converges to normal and the result of the One Way Annova test should be reliable. 

- Nevertheless, since the assumptions of the analysis of variance were broken, a non-parametric weighted Kruskal-Wallis Test was conducted to
test the equality of the means of the handshake strength variable among the studied countries. 

- The null hypothesis of equal mean across the three groups, as with the Annova test, was rejected at every level (p-value=0.00), i.e. there are
statistically significant differences in handshake strength among the countries studied.  

- A weighted Student's t-test was used separately for each pair of countries to examine which specific countries differ. This test shows that
there are statistically significant differences in average handshake strength among all the countries studied.  

```{r warning=FALSE}
#	Are there statistically significant differences between the three countries in terms of well-being? 

#  One Way ANOVA TEST
res.aov <- aov(SHARE_out$maxgrip ~ SHARE_out$country.lab, weight=SHARE_out$cciw_w7)
summary(res.aov)
#As the p-value is less than the significance level 0.05, we can conclude 
#that there are significant differences between the groups highlighted with “*" in the model summary

# check anova assumptions 
# eguality of variance
leveneTest(SHARE_out$maxgrip ~ SHARE_out$country.lab, weight=SHARE_out$cciw_w7)

#From the output above we can see that the p-value is not less than the significance level of 0.05. 
#This means that there is no evidence to suggest that the variance across groups is statistically 
#significantly different. Therefore, we can assume the homogeneity of variances in the different treatment groups.

# check normality
# Extract the residuals
aov_residuals <- residuals(object = res.aov )
# Run Kolmogorova Smirnov test
ks.test(x = aov_residuals, y='pnorm' )
plot(res.aov, 2)

##we reject Ho 0 normality of distribution


#assumption of normality doesn't met so   Kruskall Wallis test will be calculated 
#this function despite being called man whitney counts K-W test for >2 gropes to compare
weighted_mannwhitney( maxgrip ~ country.lab + cciw_w7, SHARE_out)

#Multiple pairwise-comparison between the means of groups
TukeyHSD(res.aov)

### T TESTY for each group 
# Sweden - Poland 
weights::wtd.t.test(x=SHARE_out$maxgrip[SHARE_out$country.lab=="Poland"], 
                    y=SHARE_out$maxgrip[SHARE_out$country.lab=="Sweden"], 
                    weight=SHARE_out$cciw_w7[SHARE_out$country.lab=="Poland"],
                    weighty=SHARE_out$cciw_w7[SHARE_out$country.lab=='Sweden'])


#Slovakia - Poland
weights::wtd.t.test(x=SHARE_out$maxgrip[SHARE_out$country.lab=="Slovakia"], 
                    y=SHARE_out$maxgrip[SHARE_out$country.lab=="Poland"], 
                    weight=SHARE_out$cciw_w7[SHARE_out$country.lab=="Slovakia"],
                    weighty=SHARE_out$cciw_w7[SHARE_out$country.lab=='Poland'])

# Slovakia - Sweden
weights::wtd.t.test(x=SHARE_out$maxgrip[SHARE_out$country.lab=="Slovakia"], 
                    y=SHARE_out$maxgrip[SHARE_out$country.lab=="Sweden"], 
                    weight=SHARE_out$cciw_w7[SHARE_out$country.lab=="Slovakia"],
                    weighty=SHARE_out$cciw_w7[SHARE_out$country.lab=='Sweden'])
```


- The next stage of the analysis was to examine differences between the frequency of medical consultations among the three countries studied. The
assumption of both normality of distribution of variables and homogeneity of variance were not met, so only weighted Student's  t-tests were
conducted separately for each pair of countries. 

- The tests show that all groups of countries are statistically significantly different from each other in terms of equality of the average
number of medical consultations per year.   

```{r warning=FALSE}
# Are there statistically significant differences between the three countries in terms of the frequency of visits to the doctor? 


res.aov <- aov(SHARE_out$hc002_mod ~ SHARE_out$country.lab, weight=SHARE_out$cciw_w7)

summary(res.aov)
leveneTest(SHARE_out$hc002_mod ~ SHARE_out$country.lab, weight=SHARE_out$cciw_w7)

aov_residuals <- residuals(object = res.aov )
ks.test(x = aov_residuals, y='pnorm' )
plot(res.aov, 2)
weighted_mannwhitney( maxgrip ~ country.lab + cciw_w7, SHARE_out)

TukeyHSD(res.aov)
plot(TukeyHSD(res.aov, conf.level = 0.99),las=1, col = "red")

## Here Annova's assumptions are not met so  T-Tests were performed 

# Sweden - Poland 

weights::wtd.t.test(x=SHARE_out$hc002_mod[SHARE_out$country.lab=="Poland"], 
                    y=SHARE_out$hc002_mod[SHARE_out$country.lab=="Sweden"], 
                    weight=SHARE_out$cciw_w7[SHARE_out$country.lab=="Poland"],
                    weighty=SHARE_out$cciw_w7[SHARE_out$country.lab=='Sweden'])


#Slovakia - Poland
weights::wtd.t.test(x=SHARE_out$hc002_mod[SHARE_out$country.lab=="Słowacja"], 
                    y=SHARE_out$hc002_mod[SHARE_out$country.lab=="Poland"], 
                    weight=SHARE_out$cciw_w7[SHARE_out$country.lab=="Słowacja"],
                    weighty=SHARE_out$cciw_w7[SHARE_out$country.lab=='Poland'])


# Slovakia - Sweden
weights::wtd.t.test(x=SHARE_out$hc002_mod[SHARE_out$country.lab=="Slovakia"], 
                    y=SHARE_out$hc002_mod[SHARE_out$country.lab=="Sweden"], 
                    weight=SHARE_out$cciw_w7[SHARE_out$country.lab=="Slovakia"],
                    weighty=SHARE_out$cciw_w7[SHARE_out$country.lab=='Sweden'])
```


- In addition, the relationship between the annual number of medical consultations and handshake strength separately in Poland, Sweden and
Slovakia was examined using a Chi-square test of independence. 

- For this test, the variable maxgrid (handshake strength) and the variable hc002_mod_cat (number of visits to a doctor) were aggregated to
intervals corresponding to the quantiles of the distributions of these variables. 

- The assumption of this test about the minimum size of each group (5 observations) is satisfied. 

- The null hypothesis of this test is that there is no statistical relationship between the variables under study. It was rejected for each
country (p-value=0.00), thus, there is a statistical relationship between the number of medical consultations and handshake strength of people
aged 50+ in Poland, Sweden and Slovakia. 

- Albeit, these correlations are not strong, as Cramer's V coefficient, which is a measure of the strength of the relationship between the
variables, is close to zero for each country. The strongest dependence of these variables among the countries studied can be seen in Slovakia,
where Cramer's V coefficient was 0.17.  

```{r warning=FALSE}
# Is well-being correlated with frequent doctor visits

wtd.mean(SHARE_out$maxgrip, SHARE_out$cciw_w7, na.rm=TRUE)
wtd.quantile(SHARE_out$maxgrip, SHARE_out$cciw_w7, na.rm=TRUE)


SHARE_out$maxgrip_cat <- cut(SHARE_out$maxgrip,
                         breaks=c(4, 25, 32, 42, 68),include.lowest= TRUE, 
                         labels=c( '4-24', '25-31', '32-41','42-68'))

summary(SHARE_out$maxgrip_cat)

# Poland
SHARE_out.PL <- dplyr::filter(SHARE_out,  country.lab=="Poland" )

xtp.chi2(data=SHARE_out.PL, y=maxgrip_cat, x=hc002_mod_cat, w=cciw_w7)
weighted_chisqtest(maxgrip_cat ~ hc002_mod_cat + cciw_w7, SHARE_out.PL)


# Slovakia

SHARE_out.SK <- dplyr::filter(SHARE_out,  country.lab=="Slovakia" )

xtp.chi2(data=SHARE_out.SK, y=maxgrip_cat, x=hc002_mod_cat, w=cciw_w7)
weighted_chisqtest(maxgrip_cat ~ hc002_mod_cat + cciw_w7, SHARE_out.SK)


# Sweden 
SHARE_out.SE <- dplyr::filter(SHARE_out,  country.lab=="Sweden" )

xtp.chi2(data=SHARE_out.SE, y=maxgrip_cat, x=hc002_mod_cat, w=cciw_w7)
weighted_chisqtest(maxgrip_cat ~ hc002_mod_cat + cciw_w7, SHARE_out.SE)

```


## Results of multivariate regression analysis   

- Multivariate linear regression was chosen to examine the factors that influence handshake strength.

- Three models for Poland, Sweden and Slovakia were built separately to compare how each factor affects the dependent variable in each country. 

- Control variables, in addition to the number of medical visits, were chosen for their possible effect on handshake strength among people aged
 50+. 

- The number of observations for countries varies, and some observations were removed due to missing data. 

- The control variable corresponding to occupational status was aggregated to a binary variable qualifying respondents as retired or having other
occupational status. 

- In addition, observations in which the marital status variable was a missing data were not included in the analysis so as not to distort the
interpretation of the model.  

- The F-test statistic shows that all three models are statistically significant (p-value<0.00). Although, it is worth noting that the variation
in handshake strength is best explained by the model for Sweden (the fitted coefficient of determination is 68%). For Poland, the model explains
the variation in handshake strength by only 64%, and for Slovakia by 56%.  

- The number of medical consultations turned out to be a statistically significant variable (p-value = 0.019) only for Poland and Sweden, in both
cases this variable has a negative effect, meaning that increasing the number of visits to the doctor per year by one has a negative effect on
well-being as measured by handshake strength with all other conditions unchanged. The results of the three regressions show that there are other
important determinants of good health in the elderly.  

- In all the countries studied, men had higher average handshake strength and as age increased, individuals showed weaker handshake strength,
which is also not surprising. The motor ability variable indicates difficulty in performing everyday activities (e.g., walking up stairs, taking
a shower) and as the level of difficulty increases, average handshake strength decreases for each of the three countries. 

- Only in the case of Poland did the number of years spent in the education system turn out to be a statistically significant variable: with each
additional year of education, the average handshake strength of a Polish resident increases by 0.8 measurement units with all other conditions
unchanged

- The variables material situation, self-esteem and marital status were included in the model as variables aggregated to binary variables
(dummies) to avoid collinearity. The reference point for the material situation variable is the evaluation of this situation as bad. This
variable is statistically significant for the Polish and Slovak models. It shows that the average handshake strength for those who rate their
material situation as very moderate, average or good is higher compared to those who rate their situation as bad. In the case of Slovakia, this
variable had larger coefficients which means a greater impact of moderate, medium and good material situation versus bad on well-being in people
50


- The benchmark for self-assessment of health is respondents' rating of health as excellent. In people who rate their health as either very good
or good in Sweden, handshake strength is on average lower than in people who rate their health as excellent with all other conditions unchanged,
and in Slovakia the situation is reversed. In all three countries studied, in people who rate their health as average, on average handshake
strength is lower than in people who rate their health as excellent. For those with a bad health rating, this variable is significant only for
Poland and Sweden, and maintains a negative trend: their average handshake strength is lower than for those who rate their health as excellent.

- For the marital status variable, the benchmark is the respondent's marital status of married, living together with their spouse. In Slovakia,
people in a cohabiting relationship have on average better well-being than married people living together with their spouse with all other
conditions unchanged. In the Polish and Swedish models, never-married people have, on average, weaker handshake strength than married people
living together.  

- Also included in the model was the interaction variable age*gender, which is statistically significant for each country's model, i.e. for
women,each additional year of age increases average handshake strength, i.e. inhibits the negative effect of being female on the value of the
dependent variable.  

```{r}
# Results of multivariate regression analysis 

# Clean control variable
SHARE_out$age[SHARE_out$age < 0] <- NA
SHARE_out$mar_stat[SHARE_out$mar_stat < 0] <- NA
SHARE_out$maxgrip[SHARE_out$maxgrip < 0] <- NA
SHARE_out$eduyears_mod[SHARE_out$eduyears_mod < 0] <- NA
SHARE_out$grossmotor[SHARE_out$grossmotor < 0] <- NA
SHARE_out$co007_[SHARE_out$co007_ < 0] <- NA
SHARE_out$hc002_mod[SHARE_out$hc002_mod < 0] <- NA
SHARE_out$sphus[SHARE_out$sphus < 0] <- NA

#Make categorical variable as factor 
SHARE_out$female <- as.factor(SHARE_out$female)
SHARE_out$mar_stat <- as.factor(SHARE_out$mar_stat)
SHARE_out$co007_ <- as.factor(SHARE_out$co007_)
SHARE_out$sphus <- as.factor(SHARE_out$sphus)

#Recoding
SHARE_out$mar_stat_cat <- recode(SHARE_out$mar_stat, "1=1; 2=2; 3=3; 4=4; 5=5;6=6; else=NA")
SHARE_out$Retired <- recode(SHARE_out$ep005_, "c(2, 3, 4, 5, 97)=0; 1=1; else=NA")


## Variables: 
#female
#age 
#Retired
#hc002_mod
#grossmotor
#mar_star
#co007_
#eduyears_mod


## Poland
SHARE_out.PL <- dplyr::filter(SHARE_out,
                            country.lab=="Poland" )


lm.PL <- lm(maxgrip ~ female*age+Retired+hc002_mod+grossmotor+mar_stat_cat+co007_+eduyears_mod+sphus,
           data = SHARE_out.PL,
           na.action = na.exclude)

#summary(lm.PL)


## Sweden

SHARE_out.SE <- dplyr::filter(SHARE_out,
                              country.lab=="Sweden" )
lm.SE<- lm(maxgrip ~ female*age+Retired +hc002_mod+grossmotor+mar_stat_cat+co007_+eduyears_mod+sphus,
          data = SHARE_out.SE,
          na.action = na.exclude)
#summary(lm.SE)

## Slovakia

SHARE_out.SK <- dplyr::filter(SHARE_out,
                              country.lab=="Slovakia" )
lm.SK<- lm(maxgrip ~ female*age+Retired+hc002_mod+grossmotor+mar_stat_cat+co007_+eduyears_mod+sphus,
           data = SHARE_out.SK,
           na.action = na.exclude)
#summary(lm.SK)



# Results combined 
stargazer(lm.PL, lm.SE, lm.SK, type = "text", title = "Results of linear regression analysis", 
          digits = 2, 
          column.labels = c("Poland", "Sweden", "Slovakia"), 
          dep.var.labels=c("Handshake power"),
          covariate.labels=c("Gender (female)", "Age", "Retired", "Visits to doctor", "Motor ability",
                             "Marital status: in a relationship", 'Marital status: married, living apart', 'Marital status: never married', 
                             'Marital status: divorced', 'Marital status: widower', 'Material situation: moderate', 
                             'Material situation: moderate', 'Material situation: good',
                             "Years of education", "Self-assessment of health: very good", "Self-assessment of health: good",
                             "Self-assessment of health: average", "Self-assessment of health: bad", "Women*Age", "Fixed term"), 
          out='table.txt', align=TRUE,no.space=TRUE)



#Marital status reference point: married, living together. 
#Reference point for material situation: bad.
#Reference point for self-assessed health situation: excellent. 

```

## Summary 

- A number of statistical tests conducted confirm that Poland, Sweden and Slovakia differ significantly among themselves in terms of average health grip strength, identifying well-being status, and in terms of average annual number of medical consultations. 

- Additionally, in each country studied, these variables were found to be statistically dependent among themselves. Factors influencing well-being were determined using econometric modeling. 

- First of all, the effect of the average number of visits to the doctor on handshake strength was examined. For respondents from Poland and Sweden, it turned out that an increase in the number of annual visits is associated with a decrease in the level of well-being with all other factors unchanged. 

- This gives grounds to conclude that the health status of the elderly does not depend only on the average number of medical consultations. There are other statistically significant factors affecting the handshake strength of people aged 50+ with different strength and effect in Poland, Sweden and Slovakia: age, gender, marital status, number of summers spent in education, motor ability, material situation. 

- In addition, this study underscores the existence of numerous unnecessary medical consultations, which are not only a financial burden for EU countries, but also create a problem of unavailability of doctors in critical situations for those in urgent need of medical intervention. The topic discussed needs further exploration and more thorough analysis, as it affects the level of quality of medical services in the country

